Serum Biomarkers of Nutrition Pre and Post CFTR Modulator Use in Children with Cystic Fibrosis

### 2022-04-08 Edith Zemanick/CF Nutrition/cf_nutrition.Rmd
fixed paths

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(readxl)
library(performance)
library(mcp)
library(lmerTest)
library(plotly)
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
source("~/GitHub/CF/get_genotype_class_severity.R")
```

```{r data cleaning}
# Import
df = read.csv("~/Dropbox/Work/CF/Edith Zemanick/CF Nutrition/Data_Cleaned/analysis_dataset.csv",na.strings = "")
# Only those on effective modulators
on_effective = unique(df$Patient.ID[which(df[,'Modulator'] %in% c('Kalydeco','Trikafta'))])
df = df %>% filter(Patient.ID %in% on_effective)
# Non-highy efective to "None
df$Modulator[!df$Modulator %in% c('Kalydeco','Trikafta')] = "None"
# Before or after
df$Before_After = df$Age.at.Test..Years. < df$Age.at.Start..years.
df$Before_After = factor(df$Before_After,levels = c(FALSE,TRUE),labels = c("After","Before"))
# Count number of measures 
num_values = df %>% group_by(Patient.ID,Before_After) %>% 
  summarise(n = sum(!is.na(X25OH.Vitamin.D)),.groups = "drop") %>%
  pivot_wider(id_cols = Patient.ID,names_from = Before_After,values_from = n)
num_values$`NA` = NULL
colnames(num_values)[2:3] = c("vit_measures_after","vit_measures_before")
df = left_join(df,num_values,by = "Patient.ID")
# At least two annual nutritional measurements prior to first modulator start date
two_measures = num_values$Patient.ID[which(num_values$`vit_measures_before` >= 2)]
# At least one measurement of nutritional markers and one weight/height measurement >= three months post-modulator
time_after = df %>% 
  filter(Age.at.Test..Years. >= (Age.at.Start..years. + 0.25),!is.na(X25OH.Vitamin.D)) %>%
  pull(Patient.ID) %>% unique(.)
# Filter out those without above criteria, get first modulator
df = df %>% filter(Patient.ID %in% intersect(two_measures,time_after)) %>%
  group_by(Patient.ID) %>% 
  mutate(first_mod = head(Modulator[Modulator != "None"],1)) %>% ungroup()
# Genotype group
geno = read_excel("/Users/timvigers/Dropbox/Work/CF/CF Info/Sweat Tests from Internal Database_genotype class.xlsx")
geno = geno[,c('PatientID','genoRisk')]
df = left_join(df,geno,by = c('Patient.ID' = 'PatientID'))
# Time prior to first modulator
years = df %>% group_by(Patient.ID,Before_After) %>% 
  mutate(years_to_start_max = max(abs(Age.at.Test..Years. - Age.at.Start..years.))) %>% 
  select(Patient.ID,Before_After,years_to_start_max) %>% 
  ungroup() %>% distinct(.) %>%
  pivot_wider(id_cols = Patient.ID,names_from = Before_After,values_from = years_to_start_max)
years$`NA` = NULL
colnames(years)[2:3] = c('max_years_to_start','max_years_from_start')
df = left_join(df,years,by = "Patient.ID")
df$years = df$Age.at.Test..Years. - df$Age.at.Start..years.
```

>>>>

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(readxl)
library(performance)
library(mcp)
library(lmerTest)
library(plotly)
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
}
knitr::opts_knit$set(root.dir = home_dir)
rm(home_dir)
source("~/GitHub/CF/get_genotype_class_severity.R")
```

```{r data cleaning}
# Import
df = read.csv("./Edith Zemanick/CF Nutrition/Data_Cleaned/analysis_dataset.csv",na.strings = "")
# Only those on effective modulators
on_effective = unique(df$Patient.ID[which(df[,'Modulator'] %in% c('Kalydeco','Trikafta'))])
df = df %>% filter(Patient.ID %in% on_effective)
# Non-highy efective to "None
df$Modulator[!df$Modulator %in% c('Kalydeco','Trikafta')] = "None"
# Before or after
df$Before_After = df$Age.at.Test..Years. < df$Age.at.Start..years.
df$Before_After = factor(df$Before_After,levels = c(FALSE,TRUE),labels = c("After","Before"))
# Count number of measures 
num_values = df %>% group_by(Patient.ID,Before_After) %>% 
  summarise(n = sum(!is.na(X25OH.Vitamin.D)),.groups = "drop") %>%
  pivot_wider(id_cols = Patient.ID,names_from = Before_After,values_from = n)
num_values$`NA` = NULL
colnames(num_values)[2:3] = c("vit_measures_after","vit_measures_before")
df = left_join(df,num_values,by = "Patient.ID")
# At least two annual nutritional measurements prior to first modulator start date
two_measures = num_values$Patient.ID[which(num_values$`vit_measures_before` >= 2)]
# At least one measurement of nutritional markers and one weight/height measurement >= three months post-modulator
time_after = df %>% 
  filter(Age.at.Test..Years. >= (Age.at.Start..years. + 0.25),!is.na(X25OH.Vitamin.D)) %>%
  pull(Patient.ID) %>% unique(.)
# Filter out those without above criteria, get first modulator
df = df %>% filter(Patient.ID %in% intersect(two_measures,time_after)) %>%
  group_by(Patient.ID) %>% 
  mutate(first_mod = head(Modulator[Modulator != "None"],1)) %>% ungroup()
# Genotype group
geno = read_excel("./CF Info/Sweat Tests from Internal Database_genotype class.xlsx")
geno = geno[,c('PatientID','genoRisk')]
df = left_join(df,geno,by = c('Patient.ID' = 'PatientID'))
# Time prior to first modulator
years = df %>% group_by(Patient.ID,Before_After) %>% 
  mutate(years_to_start_max = max(abs(Age.at.Test..Years. - Age.at.Start..years.))) %>% 
  select(Patient.ID,Before_After,years_to_start_max) %>% 
  ungroup() %>% distinct(.) %>%
  pivot_wider(id_cols = Patient.ID,names_from = Before_After,values_from = years_to_start_max)
years$`NA` = NULL
colnames(years)[2:3] = c('max_years_to_start','max_years_from_start')
df = left_join(df,years,by = "Patient.ID")
df$years = df$Age.at.Test..Years. - df$Age.at.Start..years.
```
